Binary Latent Representations for Efficient Ranking

### 2017-06-11 paper/paper.tex
Successfully move table.

\documentclass[sigconf]{acmart}

@@

In an on-line setting, where recommendation latency directly impacts the user experience, system designers have to work within a strict latency budget. Respecting that budget often means trading off model accuracy for speed, either via smaller and less expressive models (for example, with lower-dimensionality representations) or through aggressive use of heuristics to exclude large classes of candidate items from scoring.

The challenge of low-latency scoring leads many systems to rely on a pre-compute system, where recommendations are computed in advance and stored for serving at a later time. The primary disadvantage of such a system is its inability to react dynamically to user actions (such as new interactions that allow the recommendations to be refined) and recommendation context (location, time of day or year). Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in compute and storage resources to calculate and cache the predictions of the model.

@@

\section{Experiments}
To assess the accuracy-speed trade-offs enabled by the XNOR approach, we conduct an experiment on the Movielens 100K dataset \citep{harper2016movielens}. The dataset contains 100 thousand ratings from 943 users on 1682 movies. Despite its small size, the results on this dataset should be indicative of results that can be obtained on larger datasets, and its small size enables us to easily test hyperparameter configurations and estimate models with high latent dimensions.

\subsection{Experimental setup}
We randomly divide the dataset into a training, test, and validation set. We build real-valued and XOR models for between 32 and 4096 latent dimensions. For every latent dimensionality, we conduct a random search over the hyperparameter space, and pick optimal learning rates and minibatch sizes based on the test set. The final results are obtained from ranking interactions from the validation set.

@@

In the binary version, the 1-bit weights are packed into 32-bit integer buffers. The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result. We use the \texttt{libpopcnt} \citep{mula2016faster} library for the bit counting operations.

Both versions use 32-bit aligned input data to utilise aligned SIMD register load instructions.

The code is compiled using GCC 4.8.4 for the AVX2-enabled Broadwell architecture.

@@

\subsection{Results}
\label{sec:results}
\begin{table*}[h]
\caption{Movielens 100K results}
\label{tb:results}
\centering
\begin{tabular}{rrrrrrrr}
\toprule
Dimension &   MRR & Binary MRR & MRR ratio &   PPMS & Binary PPMS & PPMS ratio & Memory use ratio \\
\midrule
       32 & 0.096 &      0.088 &     0.919 & 89,299 &     245,644 &      2.751 &            0.091 \\
       64 & 0.098 &      0.096 &     0.973 & 47,565 &     209,486 &      4.404 &            0.062 \\
      128 & 0.108 &      0.091 &     0.843 & 20,502 &     156,404 &      7.629 &            0.047 \\
      256 & 0.109 &      0.097 &     0.884 & 10,631 &     166,500 &     15.661 &            0.039 \\
      512 & 0.112 &      0.105 &     0.938 &   5391 &     110,418 &     20.479 &            0.035 \\
     1024 & 0.114 &      0.102 &     0.897 &   2689 &      69,349 &     25.786 &            0.033 \\
     2048 & 0.111 &      0.109 &     0.984 &   1352 &      34,826 &     25.745 &            0.032 \\
     4096 & 0.111 &      0.107 &     0.961 &    666 &      15,907 &     23.856 &            0.032 \\
\bottomrule
\end{tabular}
\end{table*}
Table \ref{tb:results} summarises the results of the Movielens 100K experiment. Each row presents the best MRR (mean reciprocal rank) results for both real-valued and binary models of a given dimensionality. MRR ratio denotes the fraction of the real-valued MRR the binary model achieves for that dimensionality; PPMS (predictions per millisecond) ratio denotes how many more predictions the binary model can compute per millisecond.

@@

>>>>

\documentclass[sigchi]{acmart}

@@

In an on-line setting, where recommendation latency directly impacts the user experience, system designers have to work within a strict latency budget. Respecting that budget often means trading off model accuracy for speed, either via smaller and less expressive models (for example, with lower-dimensionality representations) or through aggressive use of heuristics to exclude large classes of candidate items from scoring. For example, approximate nearest neighbour search techniques are used at Youtube \citep{covington2016deep} to reduce the number of candidate videos that need to be scored for each user.

The challenge of low-latency scoring leads many systems to rely on a pre-compute system, where recommendations are computed in advance and stored for serving at a later time. The primary disadvantage of such a system is its inability to react dynamically to user actions (such as new interactions that allow the recommendations to be refined) and recommendation context (location, time of day or year). This leads to often complex architecture that mix fairly simple on-line models, computationally expensive offline models, and intermediate complexity near-line models where model updating is important \citep{amatriain2013big}. Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in compute and storage resources to calculate and cache the predictions of the model.

@@

\section{Experiments}
\begin{table*}[htbp]
\caption{Movielens 100K results}
\label{tb:results}
\centering
\begin{tabular}{rrrrrrrr}
\toprule
Dimension &   MRR & Binary MRR & MRR ratio &   PPMS & Binary PPMS & PPMS ratio & Memory use ratio \\
\midrule
       32 & 0.096 &      0.088 &     0.919 & 89,299 &     245,644 &      2.751 &            0.091 \\
       64 & 0.098 &      0.096 &     0.973 & 47,565 &     209,486 &      4.404 &            0.062 \\
      128 & 0.108 &      0.091 &     0.843 & 20,502 &     156,404 &      7.629 &            0.047 \\
      256 & 0.109 &      0.097 &     0.884 & 10,631 &     166,500 &     15.661 &            0.039 \\
      512 & 0.112 &      0.105 &     0.938 &   5391 &     110,418 &     20.479 &            0.035 \\
     1024 & 0.114 &      0.102 &     0.897 &   2689 &      69,349 &     25.786 &            0.033 \\
     2048 & 0.111 &      0.109 &     0.984 &   1352 &      34,826 &     25.745 &            0.032 \\
     4096 & 0.111 &      0.107 &     0.961 &    666 &      15,907 &     23.856 &            0.032 \\
\bottomrule
\end{tabular}
\end{table*}
To assess the accuracy-speed trade-offs enabled by the XNOR approach, we conduct an experiment on the Movielens 100K dataset \citep{harper2016movielens}. The dataset contains 100 thousand ratings from 943 users on 1682 movies. Despite its small size, the results on this dataset should be indicative of results that can be obtained on larger datasets, and its small size enables us to easily test hyperparameter configurations and estimate models with high latent dimensions.

\subsection{Experimental setup}
We randomly divide the dataset into a training, test, and validation set. In order to build a picture of the accuracy-speed trade-offs afforded by real-valued and binary latent models, we explicitly build models for between 32 and 4096 latent dimensions. For every latent dimensionality, we conduct a random search over the hyperparameter space, and pick optimal learning rates and minibatch sizes for each algorithm based on their performance on the test set. The final results are obtained from ranking interactions from the validation set. We use mean reciprocal rank (MRR) as a measure of ranking quality.

@@

In the binary version, the real-valued embedding parameters used in training are discarded, and the derived 1-bit weights are packed into 32-bit integer buffers. The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result. We use the \texttt{libpopcnt} \citep{mula2016faster} library for the bit counting operations.

Both versions use 32-bit aligned input data to utilise aligned SIMD register load instructions. The code is compiled using GCC 4.8.4 for the AVX2-enabled Broadwell architecture.

@@

\subsection{Results}
\label{sec:results}
Table \ref{tb:results} summarises the results of the Movielens 100K experiment. Each row presents the best MRR (mean reciprocal rank) results for both real-valued and binary models of a given dimensionality. MRR ratio denotes the fraction of the real-valued MRR the binary model achieves for that dimensionality; PPMS (predictions per millisecond) ratio denotes how many more predictions the binary model can compute per millisecond.

@@

\section{Conclusion}
We show that binary user and item representations allow recommender system designers to exploit an attractive new set of speed-accuracy trade-offs. They are straightforward to estimate, and can easily be used in any existing model estimated using stochastic gradient descent.
